\documentclass{ecai}
\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{xcolor}
\usepackage{relsize}
\usepackage{hyperref}
%%% call for papers: http://ecai2020.eu/call-for-papers/mainconference/
%%\ecaisubmission   % inserts page numbers. Use only for submission of paper.
                  % Do NOT use for camera-ready version of paper.

\newcommand{\RefPerSys}{{\textit{\textsc{RefPerSys}}}}

% defines the \rpsgitcommit command:
\input{generated-ecai2020-gitid}
\begin{document}

\title{development and architecture of {\RefPerSys}\\ {\relsize{-1}{a
      multi-threaded \textsc{ref}lective \textsc{per}sistent \textsc{sys}tem} for AI}}

\author{Basile Starynkevitch\institute{France, email: \href{mailto:basile@starynkevitch.net}{\texttt{basile@starynkevitch.net}}}
\and Abhishek Chakravarti \institute{India, email: \href{mailto:chakravarti.avishek@gmail.com}{\texttt{chakravarti.avishek@gmail.com}}}
\and  Nimesh Neema\institute{India, email: \href{mailto:nimeshneema@gmail.com}{\texttt{nimeshneema@gmail.com}}}\\
\href{http://refpersys.org/}{\texttt{refpersys.org}}}

\maketitle
\bibliographystyle{ecai}

\begin{abstract}
  %% TODO: REWRITE ABSTRACT not scientific enough.
  \textcolor{red}{REWRITE THIS}
  \textsc{\textbf{RefPerSys}} is a \textsc{\textbf{Ref}}lexive and orthogonally 
  \textsc{\textbf{Per}}sistent \textsc{\textbf{Sys}}tem (as a GPLv3+ licensed 
  free software running on Linux; it is a still unfunded research project 
  expected to last for many years, with the aim of experimenting
  with open science ideas close to Artificial General Intelligence dreams. This
  starting project is notably inspired by the works of J. Pitrat (see 
  \cite{Pitrat:2009:ArtifBeings, Pitrat:1996:FGCS, Pitrat:2009:AST, Pitrat:blog}).
\end{abstract}

\smallskip

\textbf{Topics :} Knowledge Representation and Reasoning, Machine
Learning, Multidisciplinary Topics and Applications, Agent-based and
Multi-agent Systems, Semantic Technologies.

\medskip

\section{MOTIVATIONS}
\label{sec:motivations}
\textcolor{red}{WRITE MORE OF THIS:} A symbolic AI software system
running on \textsc{Linux} needs today to manage a large amount of
information and knowledge organized as some complex graph
(a.k.a. \emph{ontologies} \cite{DeNicola:2009:OntologyBuilding},
\emph{semantic networks} \cite{VanDeRiet:1992:Ling-instr-know},
\emph{frames} \cite{Bobrow-Winograd:1977:KRL,
  greiner:1980:representation, Lenat:1983:Eurisko, Lenat:1983:theory,
  Lenat:1991:ev-cycl}) inside the computer memory. That big graph
should be orthogonally persistent \cite{Dearle:2010:orthopersist}
(like in \textsc{Bismon} \cite{Starynkevitch:2019:bismon-draft}) on
disk, and be loaded from files at startup time on mornings of worked
days, and later dumped to disk, as a set of \emph{state files}, before
normal termination when leaving office at evening. Using textual
formats on these files facilitate their management with existing
distributed version control systems such as
\href{http://git-scm.com}{\texttt{git}} or software development forges
like \href{http://gitlab.com/}{\texttt{gitlab}} and ensures some data
portability. Careful runtime generation, by metaprogramming
approaches, of \emph{C++} code (à la \textsc{Gcc Melt}
\cite{Starynkevitch-DSL2011, Starynkevitch:2009:grow,
  Starynkevitch:2007:Multistage} or \textsc{Bismon}), compiled by
\texttt{g++} into \texttt{dlopen}-ed plugins\footnote{In practice, as
shown by the \texttt{manydl.c} sample code written for J.Pitrat, on
\href{http://github.com/bstarynk/misc-basile/}{\texttt{github.com/bstarynk/misc-basile/}},
several hundreds of thousands of \texttt{*.so} plugins can be
generated then dynamically loaded by \texttt{dlopen} at runtime on
\textsc{Linux} desktops.} provides some code portability. Current
desktop computers are powerful enough to keep a large memory heap in
RAM\footnote{Notice that the C++ source of software (such as the
\textsc{Gcc} compiler) made of millions of lines of code fits entirely
in the 64 Gbytes RAM of a powerful desktop. But compiling such a code
base takes hours of computer time.}, and that entire heap can be
persisted on disk, like relational database management systems
\cite{levene:2012:relational-databases} such as
\href{http://sqlite.org/}{\textsc{Sqlite}} or \textit{NoSQL} databases
\cite{RAJ-2018-NoSQL} do. Current processors are multi-core, so
running on them a good enough multi-threaded program
\cite{butenhof:1997:programming} should be beneficial for
performance. Backtracing libraries such as I.Taylor's
\href{https://github.com/ianlancetaylor/libbacktrace}{\texttt{libbacktrace}}
facilitate introspection of the current thread's call stack, if the
\emph{C++} code has been compiled, using \texttt{g++ -O2 -g}, with
\textsc{Dwarf} debugging information. JIT-compiling libraries (such as
\href{https://gcc.gnu.org/onlinedocs/jit/}{\texttt{libgccjit}}) can
produce plugins without the need of a textual representation of some
\href{https://en.wikipedia.org/wiki/Abstract_syntax_tree}{AST} of the
code.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ack We would like to thank the referees for their comments, which
helped improve this paper considerably

\bibliography{../bib-refpersys}

%%% for a draft that should be acceptable
\begin{flushright}
  \begin{relsize}{-1}
    our draft \texttt{git commit} is \texttt{\textit{\rpsgitcommit}}
  \end{relsize}
\end{flushright}

\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% For emacs:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Local Variables: ;;
%% compile-command: "./build.sh" ;;
%% End: ;;
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
